---
title: "p8105_hw5_csf2135"
output: github_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 9,
  out.width = "90%"
)
```
### Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.


```{r, warning=FALSE, results='hide', message=FALSE}
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

df_homicides = read_csv(url(url))

df_homicides |> head()

skimr::skim(df_homicides)
```


**Describe the raw data.**

The raw data includes variables for the victim's first and last name, demographic information (sex, race, age). Each homicide has a unique identifier (`uid`) and the date the incident was reported,  the incident location (city, state, latitude, longitude), and case status (`disposition`) are provided. 


There are `r nrow(df_homicides)` observations in the data and `r ncol(df_homicides)` variables in the raw data set. Most data is non-missing, except for `r df_homicides |> filter(is.na(lat)) |> count()` instances where latitude/longitude data is missing. There are `r df_homicides |> select(uid) |> unique() |> count()` unique incidents described in the data occurring in `r df_homicides |> select(city) |> unique() |> count()` unique cities across the US. 


**Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).**
```{r, warning=FALSE, message=FALSE}
df_homicides <- df_homicides |> 
  mutate(city_state = paste(city, state, sep = ", "))

df_city_state_unsolved <- df_homicides |> 
  group_by(city_state) |> 
  summarise(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  ) |> 
  select(city_state, total_homicides, unsolved_homicides) |> 
  unique()

knitr::kable(df_city_state_unsolved, format = "markdown")
```

**For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.**
```{r,warning=FALSE, message=FALSE}
prop_unsolved = function(x) {
  
  if (is.na(x)) {
    stop("missing city name")
  } 
  
    state_unsolved_homicides <- df_city_state_unsolved  |> 
                                filter(city_state == x)  |> 
                                pull(unsolved_homicides)

    state_total_homicides <- df_city_state_unsolved  |> 
                                filter(city_state == x)  |> 
                                pull(total_homicides)
                                

    state_homicide_prop <- prop.test(state_unsolved_homicides, state_total_homicides) |> 
       broom::tidy() |>
       mutate(city_state = paste(x)) |> 
       select(city_state, estimate, `conf.high`, `conf.low`)
       
    
    return(state_homicide_prop)
}

prop_unsolved("Baltimore, MD") |> 
  knitr::kable(format = "markdown", digits = 2)
```

**Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.**
```{r}
cities <- as.vector(df_city_state_unsolved |> pull(city_state) |> unique())

df_prop_unsolved <- map_dfr(cities, prop_unsolved) 

knitr::kable(df_prop_unsolved, digits = 2)
```

**Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.**
```{r}
plot_unsolved<- df_prop_unsolved  |> 
  arrange(estimate) |> 
  ggplot(aes(x = reorder(city_state, estimate), y = estimate, color = estimate)) +
    geom_point(size = 2) +
    scale_color_viridis_c(option = "turbo") +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, alpha = 0.5) +
    labs(title = "Proportion of unsolved homicides by city", x = "City", y = "Proportion of Unsolved Homicides") +
    coord_flip() + 
    theme(
      panel.background = element_rect(fill = "grey95"), 
      axis.title = element_text(color = "black", size = 14), 
      axis.text = element_text(color = "black", size = 11), 
      title = element_text(color = "black", size = 18)
    ) 

plot_unsolved

plot_save_path <- "hw5_plots/prop_unsolved_by_state.png"

ggsave(plot_save_path, plot_unsolved, device = "png", width = 8, height = 9, units = "in", dpi = 300)

```
---
### Problem 2

**This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.**

**Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:**
```{r, message=FALSE, results='hide', warning=FALSE}
path_p2_data <- "problem2_data"

p2_csv_files <- list.files(path_p2_data, pattern = "\\.csv$", full.names = TRUE)

p2_data_frames <- map(p2_csv_files, ~ read_csv(.x) |> 
  mutate(
      subject_id = as.numeric(str_extract(.x, "(?<=_)\\d+(?=\\.csv$)")),
      arm = str_extract(.x, "([^/]+)_\\d+\\.csv$") |> str_remove("_\\d+\\.csv$")
  ))

df_p2_all <- bind_rows(p2_data_frames) |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = starts_with("week_"), 
    names_to = "week", 
    values_to = "estimate", 
    values_drop_na = TRUE
  ) |> 
  mutate(week = as.numeric(str_extract(week, "\\d+")))

```


**Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.**

```{r, message= FALSE, warning=FALSE}
plot_trial<- df_p2_all  |> 
  mutate(
    subject_arm_group = paste(subject_id, arm, sep = "_")) |> 
  ggplot(aes(x = week, y = estimate, group = subject_arm_group, color = arm)) +
  geom_line(aes(color = arm)) +
  scale_color_brewer(palette = "Dark2") + 
  labs(title = "Trial estimates overtime, by arm", x = "Study week", y = "Estimate") +
    theme(
      panel.background = element_rect(fill = "grey95"), 
      axis.title = element_text(color = "black", size = 14), 
      axis.text = element_text(color = "black", size = 11), 
      title = element_text(color = "black", size = 18), 
      legend.text = element_text(color = "black", size = 11))

plot_trial

plot_save_path2 <- "hw5_plots/trial_weekly_estimates.png"

ggsave(plot_save_path2, plot_trial, device = "png", width = 8, height = 6, units = "in", dpi = 300)
```


Over the length of the study period, the experimental group (red) consistently has higher estimates after week 2 compared to the control group. The experimental group demostrates a positive trend overtime while the control group shows a fairly flat, unchanging trendline from week 1 to week 8. 

---
### Problem 3


**When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. In this problem, you will conduct a simulation to explore power in a one-sample t-test.**

```{r}
n <- 30
sigma <- 5
mu <- 0
  
num_datasets <- as.vector(1:5000)

  
generate_datasets <- function(x) {
  dataset <- rnorm(n, mean = mu, sd = sigma)
  
  t_test_result <- t.test(dataset, mu = mu, conf.level = 0.95) |> 
    broom::tidy() |> 
    mutate(iteration = paste(x)) |> 
    select(iteration, estimate, `p.value`)

  return(t_test_result)
}

df_mu_zero_norm <- map_dfr(num_datasets, generate_datasets) 
```


**Repeat the above for μ={1,2,3,4,5,6}, and complete the following:**
```{r}
mu_list <- as.vector(0:6)

generate_datasets <- function(n, mu, sigma, x) {
  dataset <- rnorm(n, mean = mu, sd = sigma)
  
  t_test_result <- t.test(dataset, mu = 0) %>%
    broom::tidy() %>%
    mutate(iteration = x) %>%
    select(iteration, estimate, `p.value`)

  return(t_test_result)
}

dist_diff_mu <- function(mu) {
  n <- 30
  sigma <- 5
  num_datasets <- as.vector(1:5000)

  results <- map_dfr(num_datasets, ~generate_datasets(n, mu, sigma, .x)) %>%
    mutate(mu_test = paste(mu))

  return(results)
}

df_changing_mu_norm <- map_dfr(mu_list, dist_diff_mu)
```


**Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.**
```{r, message= FALSE, warning=FALSE}
perc_reject <- df_changing_mu_norm |> 
  group_by(mu_test) |> 
  filter(`p.value` < 0.05) |> 
  summarise(total_tests = n()) |> 
  mutate(perc_reject = total_tests / 5000) |> 
  arrange(mu_test) |> 
  ungroup() |> 
  ggplot(aes(x= as.numeric(mu_test), y= perc_reject)) +
    geom_point(color = "orange") +
    geom_smooth(se=FALSE, color= "orange1") +
    labs(title = "Power by true value of μ", x = "True value of μ", y = "Proportion H0 rejected") +
      theme(
        panel.background = element_rect(fill = "grey95"), 
        axis.title = element_text(color = "black", size = 14), 
        axis.text = element_text(color = "black", size = 11), 
        title = element_text(color = "black", size = 18), 
        legend.text = element_text(color = "black", size = 11))
  
perc_reject

plot_save_path3 <- "hw5_plots/perc_null_reject.png"

ggsave(plot_save_path3, perc_reject, device = "png", width = 8, height = 6, units = "in", dpi = 300)
```


Looking at the proportion of times the null was rejected (power) by ascending true effect estimates, we can see that the larger the effect size the more power the analysis of the association with have. When effect size is larger, the risk of rejecting the null when the null is true decreases. 



**Make a plot showing the average estimate of μ̂ on the y axis and the true value of μ on the x axis.**
```{r, message= FALSE, warning=FALSE}
estimate_true <- df_changing_mu_norm |> 
  group_by(mu_test) |> 
  mutate(average_estimate = mean(estimate)) |> 
  arrange(mu_test) |> 
  ungroup() |>
  ggplot(aes(x= as.numeric(mu_test), y= average_estimate)) +
    geom_point(color = "orange") +
    geom_smooth(method = "lm", se=FALSE, color= "orange1") +
    labs(title = "Average estimate of population μ by true μ ", x = "True value of μ", y = "Average estimate of μ") +
      theme(
        panel.background = element_rect(fill = "grey95"), 
        axis.title = element_text(color = "black", size = 14), 
        axis.text = element_text(color = "black", size = 11), 
        title = element_text(color = "black", size = 18), 
        legend.text = element_text(color = "black", size = 11))
  
estimate_true

plot_save_path4 <- "hw5_plots/estimated_true_corr.png"

ggsave(plot_save_path4, estimate_true, device = "png", width = 8, height = 6, units = "in", dpi = 300)
```


Given the large sample size, the average estimate of the effect size is well correlated to the true value of the effect size (orange).



**Make a second plot (or overlay on the first) the average estimate of μ̂ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. Is the sample average of μ̂ across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?**
```{r, message= FALSE, warning=FALSE}
df_estimate_true_reject <- df_changing_mu_norm |> 
  group_by(mu_test) |> 
  filter(`p.value` < 0.05) |> 
  mutate(average_estimate = mean(estimate)) |> 
  arrange(mu_test) |> 
  ungroup()

estimate_true_reject <- estimate_true + 
    geom_point(data= df_estimate_true_reject, aes(x= as.numeric(mu_test), y= average_estimate), 
               color = "green4") +
    geom_smooth(data= df_estimate_true_reject, aes(x= as.numeric(mu_test), y= average_estimate), 
                method = "lm", se=FALSE, color= "green4") 

estimate_true_reject

plot_save_path5 <- "hw5_plots/estimated_true_reject.png"

ggsave(plot_save_path5, estimate_true_reject, device = "png", width = 8, height = 6, units = "in", dpi = 300)
```


The average estimates for when the null is rejected (green) show more deviation from the true value of μ (orange) as the effect size decreases. The "significant" estimates more closely approximate the true value when the effect size is greater (>3). 




